Process data from Atlantic salmon model runs

### 2025-12-19 docs-working/01_tidying_data.qmd
Documentation updates

```{r setup, include=F, cache=F}
library(tidyverse)
library(here)
library(terra)
library(qs)
library(magrittr)
library(purrr)

@@

```{r locations}
farm_locations <- file.path(rdsi_dir, "raw_data/aquaculture-locations/all_marine_aquaculture_farms_sources_final.csv") %>% 
  read.csv() %>% 
  filter(species_group %in% c("Salmonidae fish", "General marine fish"))

@@

```{r}
nasa_dir <- "/mnt/rdsi/raw_data/NASA/GHRSST_L4_MUR_0.25_SST/"

#rename the NASA files 
#SST_files <- list.files(file.path(rdsi_dir, "raw_data/NASA/GHRSST_L4_MUR_0.25_SST"), full.names = FALSE, pattern = ".nc")
# map(SST_files, \(this_file){
#   file.rename(from = file.path(nasa_dir, this_file), to = file.path(nasa_dir, sub('^(.{4})(.*)$', '\\1_\\2',  this_file)))
# })

SST_files <- list.files(file.path(rdsi_dir, "raw_data/NASA/GHRSST_L4_MUR_0.25_SST"), full.names = TRUE, pattern = ".nc")

@@

```{r}
vuln_scores <- read.csv(file.path(rdsi_data_dir, "marine_spp_vulnerabilities/butt_spp_vuln_framework-publication/_output/vuln_gapfilled_score.csv")) %>% 
  select(vuln_gf_id, eutrophication_nutrient_pollution)

vuln_sd <- read.csv(file.path(rdsi_data_dir, "marine_spp_vulnerabilities/butt_spp_vuln_framework-publication/_output/vuln_gapfilled_sd.csv")) %>% 
  select(vuln_gf_id, eutrophication_nutrient_pollution) %>% rename(sd=eutrophication_nutrient_pollution)

vuln_tx <- read.csv(file.path(rdsi_data_dir, "marine_spp_vulnerabilities/butt_spp_vuln_framework-publication/_output/vuln_gapfilled_tx.csv"))

@@

qsave(x = full_vuln_df, file = here("data/_general_data/species_layers/vulnerabilities/marine_spp_vulnerabilities_eutrophication.qs"))
```

>>>>

```{r setup, include=F, cache=F}
library(tidyverse)
library(here)
library(terra)
library(qs)

@@

```{r locations}
farm_locations <- file.path(bigdata_path, "all_marine_aquaculture_farms_sources_final.csv") %>% 
  read.csv() %>% 
  filter(species_group %in% c("Salmonidae fish", "General marine fish"))

@@

```{r}
#rename the NASA files 
#SST_files <- list.files(file.path(rdsi_dir, "raw_data/NASA/GHRSST_L4_MUR_0.25_SST"), full.names = FALSE, pattern = ".nc")
# map(SST_files, \(this_file){
#   file.rename(from = file.path(nasa_dir, this_file), to = file.path(nasa_dir, sub('^(.{4})(.*)$', '\\1_\\2',  this_file)))
# })

SST_files <- file.path(bigdata_path, "NASA/GHRSST_L4_MUR_0.25_SST") %>% 
  list.files(full.names = TRUE, pattern = ".nc")

@@

```{r}
vuln_scores <- file.path(bigdata_path, "butt_spp_vuln_framework-publication/_output/vuln_gapfilled_score.csv") %>% 
  read.csv() %>% 
  select(vuln_gf_id, eutrophication_nutrient_pollution)

vuln_sd <- file.path(bigdata_path, "butt_spp_vuln_framework-publication/_output/vuln_gapfilled_sd.csv") %>% 
  read.csv() %>% 
  select(vuln_gf_id, eutrophication_nutrient_pollution) %>% 
  rename(sd = eutrophication_nutrient_pollution)

vuln_tx <- file.path(bigdata_path, "butt_spp_vuln_framework-publication/_output/vuln_gapfilled_tx.csv") %>% 
  read.csv()

@@

qsave(full_vuln_df, here("data/_general_data/species_layers/vulnerabilities/marine_spp_vulnerabilities_eutrophication.qs"))
```

### 2025-12-19 docs-working/02_extracting_production_locations.qmd
Documentation updates

---
title: "Extracting production locations for each species to obtain environmental forcings"
author: Richard Cottrell
date: today
editor: source
---

@@

FAO_shp <- read_sf(file.path(rdsi_dir, "raw_data/fao/FAO_AREAS_CWP_NOCOASTLINE")) %>% filter(F_LEVEL == "MAJOR") %>% select(F_CODE)

@@

@@

map_spp_to_locations_list <-

  map(.x = farm_list, .f = \(this_country_region){

@@

  })

  return(farm_locations_w_model_spp)

})

@@

```{r}

farms_w_species <- qread("data/_general_data/farm_locations/locations_w_species_fao_area.qs")

harvest_sizes <- read_csv("data/_general_data/harvest_sizes/all_harvest_sizes.csv") %>% select(-harvest_size_g)

production_cycle <- read_csv("data/_general_data/production_cycles/production_cycle.csv") %>% select(species, days)

mortality_rates <- read_csv("data/_general_data/mortality_rates/all_mortality_rates.csv")

@@

qsave(x = farms_w_stocking, file =  "data/_general_data/farm_locations/locations_w_species_fao_area_stocking.qs")

>>>>

---
title: "Extracting production locations for each species to obtain environmental forcings"
author: Richard Cottrell
date: today
execute:
  eval: false
---

@@

FAO_shp <- file.path(bigdata_path, "fao/FAO_AREAS_CWP_NOCOASTLINE") %>% 
  read_sf() %>% 
  filter(F_LEVEL == "MAJOR") %>% 
  select(F_CODE)

@@

}) %>% filter(!production_fishing_areas == "")
)
```

@@

map_spp_to_locations_list <- map(.x = farm_list, .f = \(this_country_region){

@@

  })
  return(farm_locations_w_model_spp)
})

@@

```{r}
farms_w_species <- qread("data/_general_data/farm_locations/locations_w_species_fao_area.qs") 
harvest_sizes <- read_csv("data/_general_data/harvest_sizes/all_harvest_sizes.csv") %>% select(-harvest_size_g)
production_cycle <- read_csv("data/_general_data/production_cycles/production_cycle.csv") %>% select(species, days)
mortality_rates <- read_csv("data/_general_data/mortality_rates/all_mortality_rates.csv")

@@

qsave(x = farms_w_stocking, file =  "data/_general_data/farm_locations/locations_w_species_fao_area_stocking.qs")
```

### 2025-12-19 docs-working/03_assessing_species_potential.qmd
Moving files around
Documentation updates

---
title: "Assessing species potential"
author: Richard Cottrell
date: today
editor: source
---

@@

```{r setup}
library(tidyverse)
library(here)
library(janitor)
library(countrycode)
library(rnaturalearth)

@@

```{r}
aquaculture_raw <- read_csv(file.path(rdsi_dir, "raw_data/fao/FAO_fishstat_2020/global_aquaculture_production_1950_2020.csv"))

>>>>

---
title: "Assessing species potential"
author: Richard Cottrell
date: today
execute:
  eval: false
---

@@

```{r setup, include=F, cache=F}
library(tidyverse)
library(here)
library(janitor)
library(countrycode)
library(rnaturalearth)

@@

```{r}
aquaculture_raw <- read_csv(file.path(bigdata_path, "fao/FAO_fishstat_2020/global_aquaculture_production_1950_2020.csv"))

### 2025-12-19 docs-working/04_extracting_temperatures.qmd
Moving files around
Documentation updates

---
title: "Extracting salmon farm temperature data"
author: Tormey Reimer
date: today
editor: source
---

@@

```{r setup}
#| code-summary: R setup

library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(terra)
library(qs)
library(here)
library(sf)
library(purrr)
library(furrr)
library(targets)
library(future)
library(arrow)
library(readxl)
library(units)
library(tictoc)
library(conflicted)
conflicts_prefer(dplyr::select(), dplyr::filter(), .quiet = T)

>>>>

---
title: "Extracting salmon farm temperature data"
author: Tormey Reimer
date: today
execute:
  eval: false
---

@@

```{r setup, include=F, cache=F}
#| code-summary: R setup

library(tidyverse)
library(terra)
library(qs)
library(here)
library(sf)
library(furrr)
library(future)
library(arrow)
library(readxl)
library(units)
library(tictoc)
library(conflicted)
conflicts_prefer(dplyr::select(), dplyr::filter(), .quiet = T)

### 2025-12-19 docs-working/08_data_analysis.qmd
Documentation updates

| Stat name | Unit | Description | Level |
|-------|-----|------------------------------------|
weight        | g | Mean weight of an individual fish within a farm | Individual |
dw            | g | Mean daily weight change per individual within a farm | Individual |
water_temp    | $^{\circ}$C | Mean daily temperature on farm | Farm |
T_response    | - | Temperature response of all fish within a farm (identical for all fish in a farm) | Farm |
P_excr        | g | Total excreted protein from the entire farm | Farm |
L_excr        | g | Total excreted lipid from the entire farm | Farm |
C_excr        | g | Total excreted carbohydrate from the entire farm | Farm |
P_uneat       | g | Total uneaten protein from the entire farm | Farm |
L_uneat       | g | Total uneaten lipid from the entire farm | Farm |
C_uneat       | g | Total uneaten carbohydrate from the entire farm | Farm |
food_prov     | g | Total food provided to the entire farm | Farm |
food_enc      | g | Total food encountered by each individual fish | Individual |
rel_feeding   | - | Temperature-driven relative feeding rate (identical for all fish in a farm) | Farm |
ing_pot       | g | Total ingestion potential of each (i.e. the maximum that can be eaten) | Individual |
ing_act       | g | Total actual ingestion of feed by fish in the entire farm | Farm |
E_assim       |  |  | Individual |
E_somat       |  |  | Individual |
anab          |  |  | Individual |
catab         |  |  | Individual |
O2            |  |  | Farm |
NH4           |  |  | Farm |
total_excr    | g | Total excretion from the entire farm (P+L+C) | Farm |
total_uneat   | g | Total uneaten feed from the entire farm (P+L+C) | Farm |
metab         |  |  | Individual |
biomass       | g | Mean weight of all fish within a farm (i.e. weight * population) | Farm |
: Summary of model outputs being used in this script. Note that all the outputs are daily values (i.e. value at day = t). None are summed across a production period.

>>>>

| Stat name | Unit | Description |
|-------|-----|------------------------------------|
weight        | g | Mean weight of an individual fish within a farm | Individual |
dw            | g | Mean daily weight change per individual within a farm | Individual |
water_temp    | $^{\circ}$C | Mean daily temperature on farm | Farm |
T_response    | - | Temperature response of all fish within a farm (identical for all fish in a farm) | Farm |
P_excr        | g | Total excreted protein from the entire farm | Farm |
L_excr        | g | Total excreted lipid from the entire farm | Farm |
C_excr        | g | Total excreted carbohydrate from the entire farm | Farm |
P_uneat       | g | Total uneaten protein from the entire farm | Farm |
L_uneat       | g | Total uneaten lipid from the entire farm | Farm |
C_uneat       | g | Total uneaten carbohydrate from the entire farm | Farm |
food_prov     | g | Total food provided to the entire farm | Farm |
food_enc      | g | Total food encountered by each individual fish | Individual |
rel_feeding   | - | Temperature-driven relative feeding rate (identical for all fish in a farm) | Farm |
ing_pot       | g | Total ingestion potential of each (i.e. the maximum that can be eaten) | Individual |
ing_act       | g | Total actual ingestion of feed by fish in the entire farm | Farm |
E_assim       |  |  | Individual |
E_somat       |  |  | Individual |
anab          |  |  | Individual |
catab         |  |  | Individual |
O2            |  |  | Farm |
NH4           |  |  | Farm |
total_excr    | g | Total excretion from the entire farm (P+L+C) | Farm |
total_uneat   | g | Total uneaten feed from the entire farm (P+L+C) | Farm |
metab         |  |  | Individual |
biomass       | g | Mean weight of all fish within a farm (i.e. weight * population) | Farm |
: Summary of model outputs being used in this script. Note that all the outputs are daily values (i.e. value at day = t). None are summed across a production period.
